Fast-Match: Fast and robust feature matching on large images

### 2014-12-23 paper.tex
Added comparison to geometric algorithms and reworked paper

\documentclass[10pt,twocolumn,letterpaper]{article}
\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
%\usepackage{ruler}
\usepackage{color}
\usepackage{footnote}
\usepackage{multirow} 
\usepackage{cite}
\usepackage[draft]{hyperref}
\usepackage{balance}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
%\usepackage{authblk}
\pdfminorversion=4
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

For sparse local image features many solutions have combined \emph{Ratio-Match} with various geometric constraints to improve matching. These constraints are based on assumptions regarding the transformation between the \emph{query} and \emph{target images}. A commonly used example is \emph{RANSAC} applied to feature matching where matches are chosen from a pool of candidates according to how well they approximate a global epipolar geometry \cite{fischler1981ransac,torr2000mlesac,szeliski2010}. Similar global angular and distance constraints can be used to filter a set of matches as shown in \cite{kim2008efficient,schmid1997local}. Finally the problem of feature matching can be modelled as a graph matching problem where each feature is a vertex, and edge values correspond to a geometric relation between two features as demonstrated in \cite{torresani2008feature,yarkony2010covering,cho2010reweighted}. For cases that require more flexibility such as scenes with moving elements and non-planar perspective change Pairwise constraints provide an alternative to the more rigid global constraints. Introduced by Leordeanu and Herbert \cite{leordeanu2005spectral}, pairwise constraints work by applying a geometric constraint across matches on a pairwise basis, attempting to minimize the pairwise error. This approach has later been adopted by Pang et al.\ but made scale invariant and more efficient by using a voting scheme \cite{yuan2012efficient,pang2012scale}. Similarly we can cluster matches together based on a geometric constraint and treat each cluster as a separate case \cite{cho2009feature,wu2011robust}. An interesting application of this principle is demonstrated by Chen et al.\ who iteratively use a Hough transform to cluster matches using angular constraints \cite{chen2013robust}.

@@

\section{Experimental Setup}
\label{experiments}
%
To evaluate \emph{Fast-Match} on three sets of data. We use a dataset of 3024 image pairs featuring various 3D objects at different angles and lighting conditions to measure matching performance for real life objects when lighting and perspective changes. To evaluate the performance of \emph{Fast-Match} when only part of (or none of) the scene is shared between an image pair we conduct experiments on 400 random crops with different degree of image overlap taken from 4 image pairs, creating a challenging test set where large part of the images consists of background clutter similar in nature to the subject being matched. Finally we time the algorithm on two pairs of images with high pixel counts to measure performance in terms of speed.

We compare the algorithm to the standard \emph{Ratio-Match} \cite{lowe2004sift} as well as the newer \emph{Mirror-Match} \cite{arnfred2013mirror}. We exclude the geometric algorithms discussed in related works from comparison because even the fastest cases are two to three magnitudes slower than \emph{Ratio-Match} and as such are too time consuming to use when matching features in large images.

@@

We evaluate all matching algorithms from our framework on the 3D Objects dataset by matching images at different angular intervals. For each object we pick the \emph{query} image as the image taken at 10 degrees rotation for calibration stability.  We then match this image with the same object turned an additional $\Delta$ degrees, $\Delta \in \{5, 10, \ldots, 60\}$.  For every angle interval we compare images taken under 3 different lighting conditions as provided by the dataset. We include all objects in the database for which photos at 5 degree angle intervals are available except for the ``Rooster'' and ``Sponge'' objects due to image irregularities. For each angle this means that a total of 252 image pairs are matched. We calculate the result these pairs using a weighted average based on the number of actual correspondences for each pairs. This ensures that image pairs that are more difficult to match are not contributing less to the end result because less actual matches would be found.

@@

@@

\begin{figure}[tb]
    \centering
    \includegraphics[width=1\columnwidth]{images/timings}
    \caption{The duration in seconds of different variations of
    \emph{Fast-Match} and \emph{Ratio-Match}. The top plot shows results
for two 10~MPixel images two 3~Mpixel images is used for the comparison
at the bottom. From the left, the first two results presume a precached
image while the rest do not require knowing any inputs beforehand. The
image pairs used can be seen in Figure~\ref{fig:match_example}}
    \label{fig:timings}
\end{figure}

Figure~\ref{fig:timings} demonstrate the speed of \emph{Fast-Match} on image sizes as is typically produced by modern consumer cameras and camera phones. In Figure~\ref{fig:timings} we compare the speed of the two variations of \emph{Fast-Match} to different variations of \emph{Ratio-Match} on two image pairs, one with two images of 10 megapixel and another where the same image pair has been scaled to 3 megapixels. For the larger images we see the clearest difference in performance. The retrieval variant of \emph{Fast-Match} matches the image pairs in around one second while a retrieval variant of \emph{Ratio-Match} with precomputed features spends eight seconds on the same task. This result is roughly equal to the time the general variation of \emph{Fast-Match} spends matching the two images without any pre-computations. For the nearest neighbor search we note that for large images there are substantial speed gains to be had by choosing the right algorithm for finding nearest neighbors, with Muja et al.'s Flann matcher being vastly superior to brute force and KGraph.

For the 3 megapixel picture the pattern repeats itself, although with smaller margins separating \emph{Fast-Match} and \emph{Ratio-Match}. When images are smaller, computations like obtaining the initial seed matches are comparatively more costly.

@@

We have introduced \emph{Fast-Match} in two forms, a general and a retrieval variation. The retrieval variation assumes that we know one of the images that we are matching ahead of time in order to pre-cache feature points. The general variation has no assumptions and performs the pre-caching of feature points online instead. 
We have proven that while the general variation has a computational complexity of $O(n \log n)$ with $n$ being the number of feature points, the retrieval variation will on average run in $O(n)$.
From experiments on large images we have shown that \emph{Fast-Match} can be a magnitude faster than \emph{Ratio-Match} while providing comparable results. We further compared \emph{Fast-Match} to \emph{Ratio-Match} and \emph{Mirror-Match} using images of rotated 3D objects to demonstrate that \emph{Fast-Match} outperforms the other algorithms significantly over 3024 image pairs and in most cases doubles the matching precision at similar recall rates. We show that these results hold true for matching cases featuring partial or no overlap by comparing \emph{Fast-Match} on 400 crop pairs of planar scenes.
\emph{Fast-Match} is particularly applicable in cases where we are interested in matching one image to multiple large images, in which case we can quickly pre-cache the image and use this data for each of the other images.

>>>>

\documentclass[10pt,twocolumn,letterpaper]{article}
\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
%\usepackage{ruler}
\usepackage{color}
\usepackage{footnote}
\usepackage{multirow} 
\usepackage{cite}
\usepackage[draft]{hyperref}
\usepackage{balance}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{fancyhdr}
%\usepackage{authblk}
\pdfminorversion=4
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

For sparse local image features many solutions have combined \emph{Ratio-Match} with various geometric constraints to improve matching. These constraints are based on assumptions regarding the transformation between the \emph{query} and \emph{target images}. A commonly used example is \emph{RANSAC} applied to feature matching where matches are chosen from a pool of candidates according to how well they approximate a global epipolar geometry \cite{fischler1981ransac,torr2000mlesac,szeliski2010}. Similar global angular and distance constraints can be used to filter a set of matches as shown in \cite{kim2008efficient,schmid1997local}. Finally the problem of feature matching can be modelled as a graph matching problem where each feature is a vertex, and edge values correspond to a geometric relation between two features as demonstrated in \cite{torresani2008feature,yarkony2010covering,cho2010reweighted, del2010local, puerto2011adaptive}. For cases that require more flexibility such as scenes with moving elements and non-planar perspective change Pairwise constraints provide an alternative to the more rigid global constraints. Introduced by Leordeanu and Herbert \cite{leordeanu2005spectral}, pairwise constraints work by applying a geometric constraint across matches on a pairwise basis, attempting to minimize the pairwise error. This approach has later been adopted by Pang et al.\ but made scale invariant and more efficient by using a voting scheme \cite{yuan2012efficient,pang2012scale}. Similarly we can cluster matches together based on a geometric constraint and treat each cluster as a separate case \cite{das2008event, cho2009feature, wu2011robust}. An interesting application of this principle is demonstrated by Chen et al.\ who iteratively use a Hough transform to cluster matches using angular constraints \cite{chen2013robust}.

@@

\section{Experimental Setup}
\label{experiments}
%
We use a dataset of 3024 image pairs featuring various 3D objects at different angles and lighting conditions to measure matching performance for real life objects when lighting and perspective changes. To evaluate the performance of \emph{Fast-Match} when only part of (or none of) the scene is shared between an image pair we conduct experiments on 400 random crops with different degree of image overlap taken from 4 image pairs, creating a challenging test set where large part of the images consists of background clutter similar in nature to the subject being matched. Finally we time the algorithm on two pairs of images with high pixel counts to measure performance in terms of speed.

We compare the algorithm to the standard \emph{Ratio-Match} \cite{lowe2004sift} as well as the newer \emph{Mirror-Match} \cite{arnfred2013mirror}. In addition we compare the speed of \emph{Fast-Match} to three geometric algorithms; Adaptive Multi-Affing matching (\emph{AMA}) by Suiza et al.~\cite{puerto2011adaptive}, Local shape estimation by Del bimbo et. al (Del Bimbo)\cite{del2010local} and the Isomatch matching method by Das et al\cite{das2008event}. Because of the additional time needed to run the geometric algorithms we do not include them in comparisons featuring 3D objects and partially overlapping crops. However we compare \emph{AMA} to \emph{Fast-Match} and \emph{Ratio-Match} in a more limited survey featuring the same 4 image pairs used for the partially overlapping crops since this algorithm is the fastest according to our results and outperforms \emph{Del Bimbo} according to Suiza et al.~\cite{puerto2011adaptive}.

\emph{Fast-Match} with three matching algorithms;

@@

We evaluate all matching algorithms on the 3D Objects dataset by matching images at different angular intervals. For each object we pick the \emph{query} image as the image taken at 10 degrees rotation for calibration stability.  We then match this image with the same object turned an additional $\Delta$ degrees, $\Delta \in \{5, 10, \ldots, 60\}$.  For every angle interval we compare images taken under 3 different lighting conditions as provided by the dataset. We include all objects in the database for which photos at 5 degree angle intervals are available except for the ``Rooster'' and ``Sponge'' objects due to image irregularities. For each angle this means that a total of 252 image pairs are matched. We calculate the result these pairs using a weighted average based on the number of actual correspondences for each pairs. This ensures that image pairs that are more difficult to match are not contributing less to the end result because less actual matches would be found.

@@

\begin{table}[htb]
    \caption{\emph{Fast-Match} and \emph{Ratio-Match} compared with \emph{AMA} on 4 image pairs from the Graffiti dataset by Mikolajczyk et al.\cite{mikolajczyk2005performance} as pictured in Figure~\ref{fig:patch_objects}. \emph{Fast-Match} and \emph{Ratio-Match} are evaluated with a threshold of $\tau = 0.6$ while \emph{AMA} is used with default parameters according to the reference implementation\protect\footnotemark. Precision is listed first with number of correct matches in parenthesis.}
\label{table:match}
	\centering
	\small
\begin{tabular}{r*{6}{r}}
    & \multicolumn{2}{c}{\emph{Fast-Match}} &\multicolumn{2}{c}{\emph{Ratio-Match}} & \multicolumn{2}{c}{\emph{AMA}} \\
\hline
\noalign{\smallskip}
Bark 1-4: & 0.99 & (1091) & 0.99 & (645) & 0.83 & (260)  \\
Boat 1-4: & 0.99 &(672) & 0.97 & (455) & 0.97 & (402) \\
Graf 1-3: & 0.67 &(192) & 0.72 & (141) & 0.83 & (610) \\
Wall 1-6: & 0.00 & (0) & 0.00 & (0) & 0.00 & (0) \\
\noalign{\smallskip}
\hline
\end{tabular}
\end{table}
\footnotetext{As found here: \url{http://ranger.uta.edu/~gianluca/feature\_matching/}}

% Ratio-Match
%({'bark': {'precision': 0.9984504667949431, 'recall': 645},
%  'boat': {'precision': 0.9660276729773398, 'recall': 455},
%  'graf': {'precision': 0.7157324074497083, 'recall': 141},
%  'wall': {'precision': 0.0, 'recall': 0}},

% Fast-Match
% {'bark': {'precision': 0.999999083410556, 'recall': 1091},
%  'boat': {'precision': 0.988233840832587, 'recall': 672},
%  'graf': {'precision': 0.668987216072418, 'recall': 192},
%  'wall': {'precision': 0.0, 'recall': 0}},

% AMA
% {'bark': {'precision': 0.8254, 'recall': 260},
%  'boat': {'precision': 0.9677, 'recall': 402},
%  'graf': {'precision': 0.8311, 'recall': 610},
%  'wall': {'precision': 0, 'recall': 0}})

@@

\begin{figure}[tb]
    \centering
    %
    \vspace{-2.5 mm}
    \begin{subfigure}[t]{1\columnwidth}
        \caption{3 Megapixels}
    \end{subfigure}
    \vspace{-0.5 mm}
    \begin{subfigure}[t]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/timings_medium}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.5\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/timings_geom_medium}
    \end{subfigure}%
    \vspace{-2.5 mm}
    \begin{subfigure}[t]{1\columnwidth}
        \caption{10 Megapixels}
    \end{subfigure}
    \vspace{-0.5 mm}
    \begin{subfigure}[t]{0.49\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/timings_large}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/timings_geom_large}
    \end{subfigure}%
    \vspace{1.5 mm}
    \begin{subfigure}[t]{1\columnwidth}
        \centering
        ~\includegraphics[width=1\columnwidth]{images/timings_legend}
    \end{subfigure}%

    \caption{The duration in seconds the two versions of \emph{Fast-Match} compared with \emph{Ratio-Match} and the geometric matching algorithms \emph{AMA}, \emph{Isomatch} and \emph{Del Bimbo}. The two top plots shows the duration in seconds for matching two 3~MPixel images. The bottom plots shows the duration in seconds for matching two 10~Mpixel images. For the plots on the left, the first two results of each plot presume a precached image while the rest do not require knowing any inputs beforehand. The two plots on the right shows the timings for the precached and general version of \emph{Fast-Match} identical to the plots on the left but this time compared with three geometric matching algorithms. Note that the y-scale is different on all plots. The image pairs used can be seen in Figure~\ref{fig:match_example}}
    \label{fig:timings}
\end{figure}

Table~\ref{table:match} lists the precision and number of correct matches from matching $4$ image pairs as seen in Figure~\ref{fig:patch_objects}. The geometric matching method \emph{AMA} yields better results than \emph{Fast-Match} and \emph{Ratio-Match} for the more challenging `Graf' image pair while all algorithms yield no correct matches on the `Wall' pair. For the more trivial `Boat' and `Bark' images we find \emph{Fast-Match} performing similar to \emph{Ratio-Match} in terms of precision while returning markedly more correct matches.

Figure~\ref{fig:timings} demonstrate the speed of \emph{Fast-Match} on image sizes as is typically produced by modern consumer cameras and camera phones. We compare \emph{Fast-Match} with different variations of \emph{Ratio-Match} and three geometric matching algorithms on two image pairs, one with two images of 10~Mpixels and another where the same image pair has been scaled to 3~Mpixels. The differences between \emph{Fast-Match} and \emph{Ratio-Match} are largest for larger images. The retrieval variant of \emph{Fast-Match} matches the image pairs in around one second while a retrieval variant of \emph{Ratio-Match} with precomputed features spends eight seconds on the same task. This result is roughly equal to the time the general variation of \emph{Fast-Match} spends matching the two images without any pre-computations. For the nearest neighbor search we note that for large images there are substantial speed gains to be had by choosing the right algorithm for finding nearest neighbors, with Muja et al.'s Flann matcher being vastly superior to brute force and KGraph. For the 3 megapixel picture the pattern repeats itself, although with smaller margins separating \emph{Fast-Match} and \emph{Ratio-Match}. When images are smaller, computations like obtaining the initial seed matches are comparatively more costly.

    In the plots featuring geometric matching algorithms we compare \emph{Fast-Match} with three matching algorithms; Adaptive Multi-Affing matching (AMA) by Suiza et al.~\cite{puerto2011adaptive}, Local shape estimation by Del bimbo et. al (Del Bimbo)\cite{del2010local} and the Isomatch matching method by Das et al\cite{das2008event}. For both the 3~Mpixel and 10~Mpixel images the differences between \emph{Fast-Match} and the geometric algorithms in terms of speed is vast. For the 3~Mpixel images the general \emph{Fast-Match} method is roughly $40$ times faster than the closest competitor, \emph{AMA} while it is more then $50$ times as fast as \emph{AMA} for the 10~Mpixel images. While both \emph{Isomatch} and \emph{AMA} are slow they scale well with the increased image sizes. \emph{Del Bimbo} on the other hand spends more than an hour finding matches for the 10~Mpixel image pairs.

@@

We have introduced \emph{Fast-Match} in two forms, a general and a retrieval variation. The retrieval variation assumes that we know one of the images that we are matching ahead of time in order to pre-cache feature points. The general variation has no assumptions and performs the pre-caching of feature points online instead. 
We have proven that while the general variation has a computational complexity of $O(n \log n)$ with $n$ being the number of feature points, the retrieval variation will on average run in $O(n)$.
From experiments on large images we have shown that \emph{Fast-Match} can be a magnitude faster than \emph{Ratio-Match} while providing comparable results. We further compared \emph{Fast-Match} to \emph{Ratio-Match} and \emph{Mirror-Match} using images of rotated 3D objects to demonstrate that \emph{Fast-Match} outperforms the other algorithms significantly over 3024 image pairs and in most cases doubles the matching precision at similar recall rates. We show that these results hold true for matching cases featuring partial or no overlap by comparing \emph{Fast-Match} on 400 crop pairs of planar scenes. Finally we compared \emph{Fast-Match} with the geometric matching method \emph{AMA} and note that while \emph{AMA} can yield superior matches for challenging images. However we measured \emph{Fast-Match} to be more to 40 times faster than \emph{AMA} on larger images while performing much better on image pairs that are easier to match.
\emph{Fast-Match} is particularly applicable in cases where we are interested in matching one image to multiple large images, in which case we can quickly pre-cache the image and use this data for each of the other images.
